Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-08-03 grove-paper.tex
Updates from Overleaf

@@

% \subsubsection{Wrapping Code}

% \begin{figure}
% \centering
% TODO: naive add that creates a conflict
% \begin{tikzpicture}
% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/wrapping/1.dot}};
% \path (0cm,-4cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/2.dot}};
% \path (0cm,-8cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/3.dot}};
% \path (0cm,-12cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/4.dot}};
% \path (0cm,-16cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/5.dot}};
% \end{tikzpicture}
% \caption{Wrapping}
% \label{fig:Wrapping}
% \end{figure}

% Before \marginpar{DEL?} further exploring how to merge changes from multiple users,
% we want to point out a common idiom that we call \emph{wrapping}.
% Consider the first graph in Figure~\ref{fig:Wrapping}, which
% represents the code \texttt{$\lambda$ x . x}, and suppose
% the user's cursor is in the body field of the lambda (i.e., around the variable
% reference to \texttt{x} in the body of the lambda).

% If the user requests that a plus be inserted,
% the editor may notice that there is something already in that hole.
% A naive editor could certainly add the plus by creating a new
% edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
% This would result in a conflict between the \texttt{plus} and the existing \texttt{var}
% as shown in Figure~REF:TODO.
% A better user experience
% would be interpret the request for a plus to mean to
% take whatever (if anything) is in the cursor
% and place it as a child of the plus.
% This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
% and results in the code \texttt{$\lambda$ x . x + \hole}.
% (This is similar to the behavior in other structure editors such as in CITE:TODO.)
% Translating this into edge deletes and adds results in the sequence of graphs in Figure~\ref{fig:Wrapping}.
% Note that this is a layer above

% Traditional: no relation (and indent might change)

@@

Synthetic cycles:

DIAGRAM: restoring into yourself (paste before delete?)

Organic cycles:

DIAGRAM: if cycles between "1 +" and "2 *" in previous diagrams

>>>>

\subsubsection{Wrapping Code}

TODO: move subsubsection elsewhere?

\begin{figure}
\centering
TODO: naive add that creates a conflict
\begin{tikzpicture}
\path (0,0)       node {\includegraphics[scale=0.4]{graphviz/wrapping/1.dot}};
\path (0cm,-4cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/2.dot}};
\path (0cm,-8cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/3.dot}};
\path (0cm,-12cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/4.dot}};
\path (0cm,-16cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/5.dot}};
\end{tikzpicture}
\caption{Wrapping}
\label{fig:Wrapping}
\end{figure}

Before \marginpar{DEL?} further exploring how to merge changes from multiple users,
we want to point out a common idiom that we call \emph{wrapping}.
Consider the first graph in Figure~\ref{fig:Wrapping}, which
represents the code \texttt{$\lambda$ x . x}, and suppose
the user's cursor is in the body field of the lambda (i.e., around the variable
reference to \texttt{x} in the body of the lambda).

If the user requests that a plus be inserted,
the editor may notice that there is something already in that hole.
A naive editor could certainly add the plus by creating a new
edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
This would result in a conflict between the \texttt{plus} and the existing \texttt{var}
as shown in Figure~REF:TODO.
A better user experience
would be interpret the request for a plus to mean to
take whatever (if anything) is in the cursor
and place it as a child of the plus.
This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
and results in the code \texttt{$\lambda$ x . x + \hole}.
(This is similar to the behavior in other structure editors such as in CITE:TODO.)
Translating this into edge deletes and adds results in the sequence of graphs in Figure~\ref{fig:Wrapping}.
Note that this is a layer above

Traditional: no relation (and indent might change)

@@

@@

\begin{figure}
\centering
lam x. app(\#1, \hole)
\caption{Single-User Cycles}
\label{fig:Single-User Cycles}
\end{figure}

Since we are representing programs as graphs,
we have to consider cycles.

Cycles can occur in a both single-user and multi-user contexts.

In a single user context, a cycle can appear when a user pastes
a vertex as a child of itself.
An example of this is shown in Figure~\ref{fig:Single-User Cycles},
where the user has pasted vertex~TODO as a child of vertex~TODO.

As a user interface consideration, it might be best to either
disallow such edits to at least warn users that their
edits would create a cycle.

While a single-user edits can always be checked to see
whether they create a cycle and thus the system
multi-user edits may not manifest cycles until after
the edits are merged.
there are much more practical reasons why cycles
may appear when multiple users are editing part of the code.

For example, consider the situation shown in Figure~\ref{fig:Multi-User Cycles}.
The top graph represents code that is synchronized between Alice and Bob
and represents the code \texttt{(1 * \hole) * (2 * \hole)}.
The left graph shows

TOP: *(*(1, \hole), *(2, \hole))

LEFT: *(*(1, *(2, \hole)), \hole)

RIGHT: *(\hole, *(2, *(1, \hole)))

BOTTOM: *(\hole, \hole) *(1, *(2, \#1))

\caption{Multi-User Cycles}
\label{fig:Multi-User Cycles}
\end{figure}
